---
format: pdf
editor: visual
execute:
  echo: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(fixest)
library(modelsummary)
library(gt)
library(glue)
library(tinytable)
```

```{r data-load}
POSTCODE <- 10437

dta <- readRDS("data/generated/berlin_sample.rds") %>%
	mutate(my_postcode = ifelse(
		postcode == POSTCODE, glue("Postcode: {POSTCODE}"), 
		"Rest of Berlin"
	))

n_unique_firms <- n_distinct(dta$bvdid)
n_unique_firms_my_postcode <- length(unique(
	dta$bvdid[dta$postcode == POSTCODE]
))
n_unique_firms_rest_of_berlin <- length(unique(
	dta$bvdid[dta$postcode != POSTCODE]
))

notes_sample <- glue(
	"Data are obtained from Orbis. ",
	"The sample covers the fiscal years {min(dta$year)} to ", 
	"{max(dta$year)}. It is limited to observations with postcodes ",
	"(*POSTCODE*), total assets (*TOAS*), and total equity (*SHFD*) ", 
	"available. In addition, observations are required to have ", 
	"total assets matching their sums of total equity and liabilities (*TSHF*). ",
	"Finally, observations with total equity larger than total assets are excluded. ",
	"The equity ratio is calculated by dividing total equity by total assets. ",
	"If total equity is negative it is subtracted from ",
	"total assets prior to the division ",
	"so that *equity ratio* is bounded by -1 and 1."
)
```

### Figure 1: Size and Equity Ratio over Time

#### Panel A: Size (Total Assets)

```{r Figure1PanelA, fig.align = "center", fig.height=2}
ggplot(
	dta, 
	aes(x = year, y = total_assets/1000, group = my_postcode, color=my_postcode)
) + 
	stat_summary(
		fun.data = mean_cl_normal, geom = "pointrange",
		size = 0.3, position = position_dodge(0.5)
	) +
  scale_color_brewer(
  	palette = "Set1"
  ) +
	labs(x = "", y = "Total Assets [TEUR]", color = "") +
	theme_minimal() +
		theme(legend.position = "none", panel.grid.minor = element_blank())

```

#### Panel B: Equity Ratio

```{r Figure1PanelB, fig.align = "center", fig.height=2.5}
ggplot(
	dta, 
	aes(x = year, y = equity_ratio, group = my_postcode, color=my_postcode)
) + 
	stat_summary(
		fun.data = mean_cl_normal, geom = "pointrange", 
		size = 0.3, position = position_dodge(0.5)
	) +
  scale_color_brewer(
  	palette = "Set1"
  ) +
	scale_y_continuous(labels = scales::percent) +
	labs(x = "", y = "Equity Ratio", color = "") +
	theme_minimal() +
		theme(legend.position = "bottom", panel.grid.minor = element_blank())

```

This figure compares total assets and equity ratios of Berlin firms registered under postcode `r as.character(POSTCODE)` (`r format(n_unique_firms_my_postcode, big.mark = ",")` unique firms) with the remaining population of Berlin firms (`r format(n_unique_firms_rest_of_berlin, big.mark = ",")` unique firms). `r notes_sample` 95%-Confidence intervals are plotted along with the means. Code for reproducing this figure is available at: https://github.com/joachim-gassen/rct25.

### Table 1: Size and Equity Ratio by Postcode

```{r, Table1}
myfmt <- function(x) {
	lapply(x, function (x) {
		if (x > 1) return(format(round(x), big.mark = ","))
		if (round(x, 3) == 0) return ("<0.001") 
		return(sprintf("%.3f", x))
	})
}

mod_ta <- feols(log_total_assets ~ my_postcode | year, cluster = "bvdid", data = dta)
mod_eqr <- feols(equity_ratio ~ my_postcode | year, cluster = "bvdid", data = dta)
tests <- tibble(
	`p-Value` = c(mod_ta$coeftable[1,4], mod_eqr$coeftable[1,4])
)

df <- dta %>%
	mutate(
		`Total Assets` = total_assets/1000,
		my_postcode = ifelse(
			postcode == POSTCODE,
			glue(
				"Postcode: {POSTCODE} ", 
				"(N={myfmt(length(dta$bvdid[dta$postcode == POSTCODE]))})"
			), 
			glue(
				"Rest of Berlin ",
				"(N={myfmt(length(dta$bvdid[dta$postcode != POSTCODE]))})"
			)
		)
	) %>%
	select(
		my_postcode,
		`Total Assets`, `Equity Ratio` = equity_ratio
	)

# datasummary by default uses the latex package siunitx 
# to wrap numbers and this renders \num{3,145} as 3.145 by 
# default. The option below diables the use of the siunitx package
# as our numbers already come in a pre-defined format.
options(modelsummary_format_numeric_latex = "plain")
datasummary(
	(`Total Assets` + `Equity Ratio`)~my_postcode*(Mean + SD + Median),
	df, fmt = myfmt, add_columns = tests
) %>%  theme_tt("placement", latex_float = "H")
```

This table tests for differences in total assets and equity ratios betwwen firms registered in Berlin under postcode `r as.character(POSTCODE)` (`r format(n_unique_firms_my_postcode, big.mark = ",")` unique firms) and the remaining Berlin population (`r format(n_unique_firms_rest_of_berlin, big.mark = ",")` unique firms). `r notes_sample` Total assets are measured in thousands of Euros. P-values are based on OLS regressions of both variables on a postcode indicator and a set of fiscal year fixed effects. Standard errors are clustered at the firm level. For the total assets model, the dependent variable is log-transformed to address skewness prior to estimation. Code for reproducing this table is available at: https://github.com/joachim-gassen/rct25.